The Cluster Difference Imaging Photometric Survey (CDIPS). I. Method \& TESS Light Curves from XXX Clusters

### 2019-04-06 paper/ms.tex
update paper w/ star catalog construction

@@

\title{
  The Cluster Difference Imaging Photometric Survey (CDIPS).
  I. Methods \& Light Curves from TESS Sectors 1-5
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Method}
\label{sec:method}

@@

We begin with the calibrated full frame images produced by SPOC
(\S~\ref{subsec:observations}).  We then perform a collection of
preparatory steps, including source extraction of bright stars,
astrometry using the resulting positions, and coarse simple aperture
photometry (\S~\ref{subsec:preparation}).  Using the shape values from
the initial astrometry, we select an astrometric reference frame and
transform all of the calibrated images to it.  We construct a
photometric reference by stacking a collection of frames, and then
convolve all the transformed frames to the photometric reference,
and subtract (\S~\ref{subsec:imagesubtraction}).  We perform aperture
photometry on the subtracted images using positions projected onto the
frame from Gaia DR2.  We detrend the resulting lightcurves
(\S~\ref{subsec:lcdetrending}), and then assess the resulting
lightcurves using a variety of statistics for their white noise and
red noise properties (\S~\ref{sec:results}).

@@

Second, we estimate the background of each frame.  We do this by
temporarily masking out pixels more than $2\sigma$ from the median,
and then passing a $48\times48$ median box filter over each pixel in
the image, using reflective boundary conditions. We subtract out this
smooth background estimate from each image.  This step helps mitigate
two unwanted image artifacts.  First, it removes a low-level glow
present in the corners of many images, which remains even after
flat-fielding \citep[see][\S 7.3.5]{vanderspek_2018}.  Second,
background subtraction removes scattered light patches from the Earth
and Moon, which vary in positions around the camera boresight over
each spacecraft orbit \citep[see][\S 7.3.1--7.3.4]{vanderspek_2018}.
The typical size of scattered light patches is greater than 32 pixels,
and their typical amplitude is $\sim 10\times$ the sky background
level.

@@

Finally, for frames with the \texttt{DQUALITY} bit-flag set to
32~---~corresponding to the ``momentum dumps'' described by
\citet{vanderspek_2018}~---~we mask out the entire frame.  This
removes on average a few frames per sector. Through visual inspection,
we see that the stars on these frames are extremely smeared, and are
unlikely to produce useful science data.

@@

@@

\acknowledgements
L.G.B.\ gladly acknowledges helpful discussions with
..., and is
grateful to the people who have turned TESS from an idea into reality.
%
J.N.W.\ thanks ...
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
This research has made use of the NASA Exoplanet Archive, which is
operated by the California Institute of Technology, under contract
with the National Aeronautics and Space Administration under the
Exoplanet Exploration Program.
%
This work made use of NASA's Astrophysics Data System Bibliographic
Services.
%
This research has made use of the VizieR catalogue access tool, CDS,
Strasbourg, France. The original description of the VizieR service was
published in A\&AS 143, 23.
%
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral
Agreement.
%
\newline
%
\facility{
	TESS \citep{ricker_transiting_2015},
	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}
}
%
\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{fitsh} \citep{Pal_2012},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{scipy} \citep{jones_scipy_2001}.
}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN CUSTOM SHORT-CUT COMMANDS

% BEGIN CLUSTER TABLE
%FIXME: write code that will write these values
\newcommand{\cgXVIIInstartot}{}
\newcommand{\cgXVIIInstarcdipsI}{}
\newcommand{\gaiaXVIIInstartot}{}
\newcommand{\gaiaXVIIInstarcdipsI}{}

\newcommand{\kXIIInstartot}{}
\newcommand{\kXIIInstarcdipsI}{}
\newcommand{\dXIVnstartot}{}
\newcommand{\dXIVnstarcdipsI}{}
\newcommand{\bellXVIInstartot}{}
\newcommand{\bellXVIInstarcdipsI}{}
\newcommand{\banyanXIIInstartot}{}
\newcommand{\banyanXIIInstarcdipsI}{}
\newcommand{\banyanXIInstartot}{}
\newcommand{\banyanXIInstarcdipsI}{}
\newcommand{\banyanXInstartot}{}
\newcommand{\banyanXInstarcdipsI}{}
\newcommand{\krausXIVnstartot}{}
\newcommand{\krausXIVnstarcdipsI}{}
\newcommand{\ohXVIInstartot}{}
\newcommand{\ohXVIInstarcdipsI}{}
\newcommand{\rizzutoXInstartot}{}
\newcommand{\rizzutoXInstarcdipsI}{}
\newcommand{\roserXInstartot}{}
\newcommand{\roserXInstarcdipsI}{}

\newcommand{\cgXVIIInclustertot}{}
\newcommand{\cgXVIIInclustercdipsI}{}
\newcommand{\gaiaXVIIInclustertot}{}
\newcommand{\gaiaXVIIInclustercdipsI}{}

\newcommand{\kXIIInclustertot}{}
\newcommand{\kXIIInclustercdipsI}{}
\newcommand{\dXIVnclustertot}{}
\newcommand{\dXIVnclustercdipsI}{}
\newcommand{\bellXVIInclustertot}{}
\newcommand{\bellXVIInclustercdipsI}{}
\newcommand{\banyanXIIInclustertot}{}
\newcommand{\banyanXIIInclustercdipsI}{}
\newcommand{\banyanXIInclustertot}{}
\newcommand{\banyanXIInclustercdipsI}{}
\newcommand{\banyanXInclustertot}{}
\newcommand{\banyanXInclustercdipsI}{}
\newcommand{\krausXIVnclustertot}{}
\newcommand{\krausXIVnclustercdipsI}{}
\newcommand{\ohXVIInclustertot}{}
\newcommand{\ohXVIInclustercdipsI}{}
\newcommand{\rizzutoXInclustertot}{}
\newcommand{\rizzutoXInclustercdipsI}{}
\newcommand{\roserXInclustertot}{}
\newcommand{\roserXInclustercdipsI}{}
% END CLUSTER TABLE

% END CUSTOM SHORT-CUT COMMANDS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%

@@

\title{
  The Cluster Difference Imaging Photometric Survey (CDIPS).
  I. Method \& TESS Light Curves from XXX Clusters
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Method: Photometry}
\label{sec:method}

@@

We begin with the calibrated full frame images produced by SPOC
(\S~\ref{subsec:observations}).  We then perform a collection of
preparatory steps, including source extraction of bright stars,
astrometry using the resulting positions, and coarse simple aperture
photometry (\S~\ref{subsec:preparation}).  Using the shape values from
the initial astrometry, we select an astrometric reference frame and
transform all of the calibrated images to it.  We construct a
photometric reference by stacking a collection of frames, and then
convolve all the transformed frames to the photometric reference,
and subtract (\S~\ref{subsec:imagesubtraction}).  We perform aperture
photometry on the subtracted images using positions projected onto the
frame from Gaia DR2.  We detrend the resulting lightcurves
(\S~\ref{subsec:lcdetrending}).
We then select stars that we believe are ``cluster members'',
roughly X\% of the overall dataset (\S~\ref{sec:starselection}).
Focusing on the lightcurves of these stars, their white noise and red
noise properties, and point out a few interesting cases of
variability (\S~\ref{sec:results}).

@@

We then estimate the background of each frame.  We do this by
temporarily masking out pixels more than $2\sigma$ from the median,
and then passing a $48\times48$ median box filter over each pixel in
the image, using reflective boundary conditions. We subtract out this
smooth background estimate from each image.  This step helps mitigate
two unwanted image artifacts.  First, it removes a low-level glow
present in the corners of many images, which remains even after
flat-fielding \citep[see][\S 7.3.5]{vanderspek_2018}.  Second,
background subtraction removes scattered light patches from the Earth
and Moon, which vary in positions around the camera boresight over
each spacecraft orbit \citep[see][\S 7.3.1--7.3.4]{vanderspek_2018}.
The typical size of scattered light patches is greater than 48 pixels,
and their typical amplitude is $\sim 10\times$ the sky background
level.

@@

Finally, for frames with the \texttt{DQUALITY} bit-flag corresponding
to the ``momentum dumps'' and ``coarse pointing modes'' described by
\citet{vanderspek_2018}, we mask out the entire frame.  This removes
on average a few frames per sector. Through visual inspection, we see
that the stars on these frames are extremely smeared, and are unlikely
to produce useful science data.
In addition, we use the sector-specific data release
notes\footnote{\url{ archive.stsci.edu/tess/tess_drn.html}} to
identify further  times with anomalously poor spacecraft performance,
which we omit from consideration. For Sectors 1 through 5, these
include coarse pointing windows from orbits 10, 13, 14, and an
instrumetal anomaly window in orbit 15.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Method: Star Selection}
\label{sec:starselection}

A major motivator for the CDIPS project is to increase the number of
cluster stars for which continuous photometric timeseries are
available, and thus facilitate studies of exoplanetary and stellar
processes across different times and stellar environments.

An essential aspect of the project is therefore to define a {\it
cluster star sample}, consisting of stars that are thought to be
members of clusters.

Performing a homogeneous membership determination for every known
cluster is beyond the scope of this work.  Instead, we collect various
cluster membership catalogs from the literature.  We then use them to
identify cluster members for which we have already made lightcurves
from the TESS images.  Our goal in selecting stars is completeness,
rather than precision.  If there has been a claim in the literature
that a star should be considered as a cluster member, we aim to
provide a lightcurve for that star.

Table~N describes the cluster membership catalogs we have used to
identify candidate cluster members.  Since our photometric reduction
is based on positions reported in {\it Gaia DR2}, each of our
lightcurve sources by default is associated with a Gaia identifier.
Correspondingly, the cluster membership determinations performed by
\citet{cantat-gaudin_gaia_2018} and \citet{gaia_hr_2018} are the
easiest case for us to merge against our lightcurve database.
For the TESS Sectors 1--5 lightcurves produced in this study, this
yielded X and Y lightcurves respectively.

%FIXME: make table

\subsection{Open clusters (OCs)}
\label{subsec:oc}

\paragraph{{\it Gaia}-derived OC memberships}
At the time of writing, two relatively large, homogeneous cluster
memberships studies had been performed using {\it Gaia}-DR2: those by
\citet{cantat-gaudin_gaia_2018} and \citet{gaia_hr_2018}.

\citet{cantat-gaudin_gaia_2018} used an unsupervised membership
assignment algorithm to identify clusters in the three-dimensional
astrometric space of proper motion and parallax. They used {\it Gaia}
photometry and radial velocities to then verify the claimed
membership properties.  From their Table~2, we collect 401{,}448
cluster members, in 1229 clusters, down to a limiting magnitude of
$G=18$.

\citet{gaia_hr_2018} reported memberships for stars in a smaller, more
select group of well-studied open clusters. From their Table~A1, we
collect 40{,}903 cluster members, in 41 open clusters, mostly within
$500\,{\rm pc}$. While this work also included memberships for
globular clusters, we omitted these from consideration.

\paragraph{Pre-{\it Gaia} OC memberships}
\citet{Kharchenko_et_al_2013} used proper motions calculated in PPMXL
\citep[][a combination of USNO-B1{.}0 and 2MASS
astrometry]{roeser_ppmxl_2010} and near-infrared photometry from 2MASS
\citep{skrutskie_tmass_2006} and reported the existence of 2859 open
clusters and stellar associations.
We selected their ``$1\sigma$'' members according to the
combined photometric, kinematic, and spatial criteria described by
\citet{kharchenko_global_2012}.  Then, to obtain {\it Gaia}-DR2 source
identifiers for the members, we performed a crossmatch for {\it
Gaia}-DR2 sources within 5 arcseconds of the listed positions.  As an
additional constraint, we used the 2MASS photometry to predict the
$G$-band magnitudes\footnote{See
\url{https://gea.esac.esa.int/archive/documentation/GDR2/Data_processing/chap_cu5pho/sec_cu5pho_calibr/ssec_cu5pho_PhotTransf.html},
online, \texttt{2019-03-29}, or \citet{carrasco_gaia_2016}}, and
required that the measured $G$-magnitude fall within 2 magnitudes of
the predicted $G$-magnitude.  If multiple neighbors matched the
position and magnitude constraints, we took the nearest spatial
neighbor as the match.  From 373{,}226 stars, this yielded a unique
best neighbor for 352{,}332 stars (94.4\% of the sample), and a choice
between two neighbors for 17{,}774 stars.

The second (non-{\it Gaia} derived) open cluster membership catalog we
used was the \citet{dias_proper_2014} catalog, which was based on
UCAC4 proper motions.
From their 1805 reported open clusters, we selected sources with
quoted membership probability above 50\%.
To obtain {\it Gaia}-DR2 source identifiers for the members, we
performed a similar crossmatch as before, looking for sources within 5
arcseconds of the listed positions, and within $\pm$2 $G$-band
magnitudes of the prediction.
From 2{,}034{,}269 stars, this yielded a unique
best neighbor for 1{,}828{,}630 stars (89.9\% of the sample), and a choice
between two neighbors for 8.7\% of the remaining sample.

\begin{figure*}[!ht]
  \gridline{\fig{f5a.png}{0.95\textwidth}{}}
  \vspace{-0.8cm}
  \gridline{\fig{f5b.png}{0.95\textwidth}{}}
  \caption{
      {\it Top.} Cross-match statistics from \cite{Kharchenko_et_al_2013} {\it 
      vs.} Gaia-DR2.
      {\it Bottom.} Ditto, for \cite{dias_proper_2014} {\it vs.} Gaia-DR2.
  }
  \label{fig:xmatch_info}
\end{figure*}

The distributions of various cross-matching statistics are shown in
Figure~\ref{fig:xmatch_info}.  The distances between matches is
typically below 1 arcsecond.  The Dias catalog shows somewhat stronger
crowding effects at the faint end compared to the Kharchenko catalog.
The Kharchenko catalog also has a more lop-sided distribution of true
{\it vs.} predicted $G$-band magnitudes.

\subsection{Moving groups and stellar associations}
\label{subsec:mg}

Stars, moving groups and stellar associations are of interest for
similar reasons as stars in open clusters.  Though fewer stars
are known to exist in moving groups, they are of particular interest
because moving groups are less crowded than open clusters, and are
often closer to the Sun.

We obtained Gaia DR2 identifiers from the results of the following
studies:
\citet{gagne_banyan_XI_2018},
\citet{gagne_banyan_XII_2018},
\citet{gagne_banyan_XIII_2018},
\citet{kraus_tucanahor_2014},
\citet{roser_deep_2011}, % OC, not MG...
\citet{bell_32ori_2017},
\citet{rizzuto_multidimensional_2011},
and \citet{oh_comoving_2017}, The methods applied in these studies
vary from kinematic analyses, to astrometric analyses included
Gaia-DR1 parallaxes, to photometric searches for infrared excesses, to
spectroscopic studies including RVs, H$\alpha$
emission, and Li absorption.

For the Gagne et al{.} catalogs, we searched the Gaia-DR2 archive for
sources within 10 arcseconds of the listed positions.  If Gagne et
al{.} gave a proper motion, we required that the sign of each the Gaia
proper motion components match that of the Gagne values (the stated
proper motion uncertainties seemed to have been underestimated).  We
also imposed a $G<18$ cut on any putative matches.  Of 3012 moving
group members collected from the three combined Gagne et al{.}
catalogs, we found 2702 matches.

The \citet{kraus_tucanahor_2014}, \citet{roser_deep_2011}, and
\citet{bell_32ori_2017} studies reported members in Tucana-Horologium,
the Hyades, and 32$\,$Ori respectively.  Applying the same procedure as
for the Gagne catalogs gave 187, 684, and 119 best-neighbors
respectively, compared to 205, 724, and 141 initially reported
members.  Note that \citet{kraus_tucanahor_2014} found that only
$\sim$70\% of their listed members have spectroscopic indicators
consistent with their membership in Tucana-Horologium.

\citet{rizzuto_multidimensional_2011} also focused on a single moving
group: the Sco OB2 association. We used their reported Hipparcos
identifiers, and matched against the {\it Gaia} archive's
\texttt{hipparcos2_best_neighbour} table, which gave 319
nearest-neigbor stars from 436 candidate members.

Finally, \citet{oh_comoving_2017} searched for comoving stars in the
$\approx$2 million stars that overlapped between Tycho-2 and {\it
Gaia}-DR1.  They found many wide binaries, and also identified a large
number of comoving groups.  We chose the 2{,}134 stars that they
reported were in groups with sizes of at least 3 stars.  Using their
{\it Gaia}-DR1 source identifiers, we matched against the {\it Gaia}
archive's \texttt{dr1_neighbourhood} table, which gave 1{,}881
nearest-neigbor stars in groups of at least three stars
\citep{marrese_gaia_2019}.

\subsection{Summary of selected stars}

After collecting the results of all the studies described above, we
merged them into a single table. We queried the
\texttt{gaiadr2.gaia_source} table to retrieve their photometric $G$,
$G_{\rm R_p}$, and $G_{\rm B_p}$ magnitudes, as well as their
five astrometric parameters $(\alpha, \delta, \mu_\alpha, \mu_\delta,
\pi)$.
We then imposed that $G_{\rm R_p} < 16$.
The concatenated results are given in Table~N.

All told, there are 1{,}308{,}706 unique stars, from 12 distinct
membership catalogs.
Stars that are reported in multiple catalogs have their reference
information from each available catalog concatenated.

@@

\acknowledgements
L.G.B.\ gladly acknowledges helpful discussions with
..., and is
grateful to the people who have turned TESS from an idea into reality.
%
J.N.W.\ thanks ...
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
This research has made use of the NASA Exoplanet Archive, which is
operated by the California Institute of Technology, under contract
with the National Aeronautics and Space Administration under the
Exoplanet Exploration Program.
%
This work made use of NASA's Astrophysics Data System Bibliographic
Services.
%
This research has made use of the VizieR catalogue access tool, CDS,
Strasbourg, France. The original description of the VizieR service was
published in A\&AS 143, 23.
%
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral
Agreement.
%
\newline
%
\facility{
	TESS \citep{ricker_transiting_2015},
	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018},
	2MASS \citep{skrutskie_tmass_2006}
}
%
\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{fitsh} \citep{Pal_2012},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{scipy} \citep{jones_scipy_2001}.
}
